%---------------------------------------------------------
\section*{Extras}
%---------------------------------------------------------

%------------------------------------------------
\begin{frame}{Why We Need \BVCAS}
\begin{itemize}
  \item In Byzantine settings, simple registers are unsafe:
    \begin{itemize}
      \item A correct value can be overwritten by a faulty process.
      \item Leads to inconsistent states and broken agreement.
    \end{itemize}
  \item Sticky bits resist corruption but support only binary values.
  \item \BVCAS extends the idea to multi-valued consensus.
\end{itemize}
\end{frame}



\begin{frame}{Why Assumptions are Reasonable}
\begin{itemize}
  \item RR scheduling: baseline GPU behavior~\footfullcite{Jeon:springer:2025}.
  \item \Fwrite{} has constant cost (fits in one phase).
  \item \Fview{} may span multiple phases, consistent with GPU memory latency.
  \item Restricted access via objects is common for fault-tolerant designs.
\end{itemize}
\end{frame}

\begin{frame}{Weak Byzantine Consensus (WBC)}
\begin{definition}[\wbc] 
	Given $n$ processes, up to $f<n$ faulty, and correct values $V_c$, the following must hold:
	\begin{itemize}
		\item \textbf{Termination: } Every correct process must eventually decide.
		\item \textbf{Agreement: } All correct processes decide the same value.
	\end{itemize}
\end{definition}
\begin{itemize}
  \item Validity condition is dropped.
  \item Even if all correct processes propose the same value, the decision could be different.
\end{itemize}
\end{frame}

%------------------------------------------------
\begin{frame}{Common Value Byzantine Consensus (CVBC)}
\begin{theorem}[\cvbc] 
Proposed solution solves solves the \cvbc problem within $\Theta(n^2/p^2r)$ phases, for $f< \frac{n}{2p}$.
\end{theorem}
\end{frame}

